---
title: "Introduction to greenfeedr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to greenfeedr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

When working with GreenFeed system you must:

* Retrieve data from each GreenFeed unit
* Check animal visitation
* Filtering and process data for further analysis

The greenfeedr makes these steps fast and easy:

* It provides functions for downloading, reporting, and processing GreenFeed data

This document introduces you to greenfeedr's set of tools, and shows you how to apply them to data frames.

## Download data with `get_gfdata()`

The first step in working with the GreenFeed system is to retrieve your data. This can be done manually by logging into the GreenFeed web interface [C-Lock Inc.](https://ext.c-lockinc.com/greenfeed.php)

However, the easiest way to retrieve your daily data is via the API using the `get_gfdata()` function. This function automates the data retrieval process, allowing you to specify user and password, units, and date range to download the data directly into your computer.


## GreenFeed data

To explore the basic data manipulation functions of greenfeedr, we'll use dairy cow data from a experiment with 32 cows.

```{r data}
data <- readr::read_csv(system.file("extdata", "StudyName_GFdata.csv", package = "greenfeedr"), show_col_types = FALSE)
dim(data)
data
```



## Report data with `report_gfdata()`







